The Splashback Radius of Planck-SZ clusters

### 2019-02-13 Paper_v2/Reply_to_referee.txt

We thank the referee for the very detailed and useful report. We have made the
changes based on the referee's detailed suggestions, and these changes should
have improved the readability of the paper. We hope the revised manuscript is
in an acceptable format. In our response below, the referees comments have been
quoted, and our response indicated by Reply.

Please let us know if there are any further concerns.

Cheers,
Dominik and Surhud

> Referee Report
> 
> In previous work, the splashback radius of dark matter halos was detected by
> optically selecting galaxy clusters and measuring the galaxy (or lensing mass)
> distribution around them. In this paper, the authors propose an entirely new
> method, namely selecting clusters in Planck SZ maps. The main advantage is to
> avoid the troubling systematics in optical selection that have recently been
> discovered.
> 
> The paper represents a highly valuable contribution to the field, and will
> certainly spark many follow-up studies using other datasets. Overall, the paper
> is already in good shape: the results and the method by which they were
> obtained are clear and sensible, and the writing is (mostly) clear. I do,
> however, fear that the current version is somewhat cluttered with too many
> figures and tables that might confuse the reader and lessen the paper's impact.
> Moreover, some of the writing assumes a fair amount of previous knowledge of
> the subject, meaning that readers unfamiliar with splashback might be turned
> off.
> 
> I will be happy to recommend the paper for publication once the major points
> below have been addressed. I emphasize that many of my suggestions are just
> that: writing, figure design, and the flow of a paper are, to a large extent, a
> matter of taste. I hope that you will find my suggestions helpful.
> 
> Congratulations on this really important result.

Reply: We thank the referee for a careful reading of the manuscript and the
detailed suggestions which have helped improve the paper.

> 
> 
> 1) The main points of the paper are somewhat hidden in a wealth of 16 figures
> and 6 tables, some of which seem detail-oriented and make the paper harder to
> digest than it needs to be. The suggestions below would drastically reduce the
> number of figures, without losing any essential information.
> 
> a) Figure 1: It is not clear why the sky maps of the selected clusters are
> buried in the appendix while the mass-redshift distribution is shown in Figure
> 1. Both figures are relatively technical. My recommendation would be to shift
> Figure 1 into Appendix A, which could be renamed from "Sky maps" to "Properties
> of the SZ sample" or something like that.

Reply: Our original motivation to keep Figure 1 in the beginning was to show the
average redshift of the sample and the mass range, since that goes in our
theoretical expectation. However since this is indicated in the Table, we have
now moved Figure 1 to the appendix based on the referee's suggestion.

> 
> b) Figures 2 and 3 show very similar things, namely the 2D and 3D density
> profiles. However, because the figures are split and because the panels come
> out much larger in Figure 3, it is not immediately apparent that the actual
> measured results are shown in Figure 2. I have a somewhat radical suggestion: I
> would combine both figures by stacking the panels vertically instead of
> horizontally, i.e., combining the 2D and 3D profiles into one column of five
> panels that shows all aspects of the modeling (data, inferred 2D, 2D slope, 3D
> etc). Then, the three samples would be three columns that could easily be fit
> next to each other, particularly as the y-labels would not need to be repeated
> in each column. The disadvantage is that it would become harder to compare the
> location of Rsp in the three samples, but as stated in the text, they all agree
> anyway. The advantage would be that one could directly compare the inferred 2D
> and 3D radii -- to me, that is more informative. Most importantly, it would be
> very clear that the combined figure represents the main result of the paper.
> The same could be done for Figures 6 and 7 which, again, appear somewhat
> disconnected to the reader even though they are highly connected.

Reply: We thank the referee for this suggestion. We have now combined Figure 2
and Figure 3 in the original manuscript into one. However, since the upper two
panels have the projected radius as the abscissa while the bottom three have
the 3-dimensional radius instead, we have included the labels at the bottom of
row 2, and row 5 separately. {\bf TBD: We have also included a label Rsp^{\rm
2d} and Rsp^{\rm 3d} to indicate the two and three dimensional locations of the
steepest slopes.} We have similarly combined Figure 6 and 7 into the new Figure
4.

> 
> c) Tables 3 and 4 contain so little data that they could easily be put into
> words and added to the text.

Reply: We agree and have described this table in the text instead.

> 
> d) Figures 4 and 5 could easily be combined into one figure, perhaps with
> clear, large labels inside the panels explaining what the different
> distributions represent.

Reply: We apologize that the right panel in Figure 4 was inadvertently included
in the original manuscript. We have now removed that part and combined Figure 4
and 5 into a single one and included large labels inside the panels explaining
what each of the distribution corresponds to.

> 
> e) Figures 8 and 9 could be combined by showing the clusters on top of the
> masks. That way, the masks would be more useful because the reader could
> immediately see how many clusters are masked out.

Reply: We agree and have changed the figure as suggested.

> 
> f) Showing the corner plots in Figures 10 to 14 is laudable in terms of
> absolute completeness of the information presented, but frankly, I am not sure
> they are necessary. In particular, Figures 10-12 are virtually impossible to
> distinguish by eye, and Figures 13-14 are difficult to digest. I had to stare
> at these figures quite hard to derive any meaningful takeaway messages. My
> suggestions would be to omit all of these figures except perhaps one for one of
> the PanStarrs samples. That figure could then even be put into the main body of
> the paper instead of an appendix.

Reply: We agree with the referee's suggestion and have included one of the
corner plots in the main manuscript and the others have been removed from the
appendix. We have also included some messages that we want the reader to take
from these plots in the text describing this figure.

> 
> g) I am not sure that figures 15 and 16 are necessary. I found the written
> explanation of the checks on color scatter perfectly adequate and convincing
> without the figures.

Reply: We agree that the figure XX corresponding to the error as a function of
magnitude may not be necessary, but we would like to show the color cut we use
for reproducibility purposes. We hope the referee will be ok with this.

> 
> 2) Throughout the paper, I found that the writing supposed the reader to
> possess a high level of knowledge about the specifics of the splashback radius
> and its detection. It would be good to add some gentle explanations as well as
> smoother transitions between different sections and arguments. I give some
> examples below.
> 
> a) In the abstract, an introductory sentence would help. What has been done so
> far, why move from optical to SZ selection? A comment to this effect would also
> be good in Section 2.1, some readers are probably confused as to the purpose of
> the SZ sample (which is used only to find clusters but not to constrain their
> profiles).

Reply: We have now included sentences to this effect in the {\bf abstract},
the introduction as well as just before Section 2.1.

> 
> b) Throughout the paper, it is implicitly assumed that the definition of the
> splashback radius is the radius where the 3D slope of density is steepest, but
> that is only one of many definitions. It was defined this way in More+15 as the
> most easily measurable quantity, but one can also use the radius of the
> outermost shell in theoretical models (Adhikari+14, Shi 16), non-spherical
> shells (Mansfield+17), or the apocenter radii of particles (Diemer 17). To be
> clear, this comment is not intended as a sneaky citation request. I do,
> however, feel that there is already a lot of confusion as to "what the
> splashback radius actually is", and any splashback paper should try to
> alleviate this confusion.

Reply: We would like to apologize for missing these citations. We have now
pointed out these alternative ways of defining the splashback radius in the
Methods section, along with the relevant citations.

> 
> c) It is assumed that the reader understands the connection between the
> limiting magnitude of galaxy samples and dynamical friction, but that
> connection is far from obvious! Please explain that dynamical friction can
> reduce Rsp, that this effect is mass-dependent, and that selecting more massive
> galaxy samples can thus put limits on the effect. In this context, it would
> also be good to cite Adhikari+16 who measure and model the magnitude of
> dynamical friction in SDSS.

Reply: We have now added comments relevant to dynamical friction effects in the
introduction itself now.

> 
> d) One common concern is that the profile fitting function might introduce a
> splashback scale simply due to its functional form. This was shown not to be
> the case in Baxter+ and Umetsu & Diemer 17, it would be good to explain that.

Reply: We have now added the relevant comments in the methods section.

> 
> e) The corner plots in Figures 10-14, whether they remain in the paper or not,
> would benefit from some explanation. These plots are hard to read: which
> parameters are well-constrained, which are not? Which are dominated by priors?
> Are there important degeneracies? What does all of this mean for the inferred
> splashback radii?

Reply: We have now kept only one of the corner plots, and have written the main
takeaways from the plot in the Results section of the text.

> 
> 3) Some results seem incompatible with results in Baxter+17, and (if real)
> those differences should be discussed.

{\bf Still to be done?}

> a) The right panel of Figure 4 seems puzzling, should these slopes not be much
> steeper? Compare to Figure 4 in Baxter+17.

Reply: This was an erroneous plot, we apologize for not noticing this before,
and thank the referee for noticing it.

> 
> b) In Section 4, you explain that the splashback radius seen in the blue
> population may be due to contamination of the blue sample, but the conclusion
> remains a little unclear in the main text. I would refer to Baxter+17 Figure 5,
> saying that they found no splashback in the blue population, and make a firm
> statement on whether there is any evidence that this is wrong (given the issues
> with contamination).  -refer to baxter and explain why we find something
> different
> 
> 4) I have some smaller comments and questions regarding the text and figures.
> 
> a) In Table 2, it looks like "'" were used instead of "." in large numbers.

Reply: We have corrected this.

> 
> b) The derivatives in figures 2, 3 etc look somewhat noisy, even though the
> profile measure is smooth as it is inferred from the profile fitting function.
> My expectation would be that the derivatives should look smooth too?

Reply: Not corrected since this would require evaluating function on many more
points -> large computational cost

> c) Section 4.1 is the only sub-section of Section 4, and thus makes no sense. I
> would either remove the 4.1 section title or add more sub-sections.

Reply: Done

> d) In Section 4, you state that you are using an "average concentration-mass
> relation" to convert M500c to M200m. Which model of the concentration-mass
> relation was used?

Reply: We now mention the model of concentration-mass relation we use.

> 
> e) In Figures 4 and 5, y-labels are missing (even though they are fairly
> obvious). Moreover, it would be good to increase the size of the inset labels
> that state what each panel is showing.

Reply: We have added the y-labels and increased the size of the inset labels.

> 
> f) Figure 8 looks like a low-resolution pixel graphic. Is there a reason not to
> show it as a vector graphic?

Reply: We have now edited this.

> g) Table 6 would be more useful if it showed the Rsp values computed without
> miscentering for comparison.

### 2019-02-13 Paper_v2/RSP_paper_new.tex
edits to paper

\begin{abstract}
We present evidence for the existence of the splashback radius in
galaxy clusters selected using the Sunyaev-Zeldovich effect.  We show
that the deprojected cross-correlation of galaxy clusters found in the
Planck survey with galaxies detected photometrically in the Pan-STARRS
survey, shows a sharp steepening feature (a logarithmic slope steeper
than $-3$), which we associate with the splashback radius. We infer
the three-dimensional splashback radius for the SZ cluster sample to
be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$ $\mpch$, where the cluster sample
has an average halo mass $M_{\rm 500c}=3.0\times10^{14}\msunh$ at an
average redshift of $z=0.18$. The inferred value of the splashback
radius appears consistent with the expected location for dark matter halos
in the standard cold dark matter paradigm. However, given the limited
precision of our measurements, we cannot conclusively rule out the
smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius
does not depend upon the galaxy magnitude for galaxies fainter than
$M_i-5\log h=-19.44$, and is present at a consistent location in
galaxy populations divided by color. The presence of the splashback
radius in the star-forming galaxy population could potentially be used to put
lower limits on the quenching timescales for galaxies. We can marginally rule
out the contamination of the star-forming galaxy sample by quenched galaxies,
but the results would need further verification with deeper datasets.

@@

\textbf{To avoid optical cluster selection effects, there have been some
previous efforts to use X-ray selected cluster samples to look for the
splashback radius using the weak lensing signal \citep{umetsu2017lensing,
contigiani2018weak}}. However stacking issues and the low signal-to-noise
ratio remains a significant hurdle for both of them. \textbf{The use of weak
lensing data in these studies could get around the biased nature of the galaxy
distribution by directly probing the dark matter distribution. Along those
lines,} \citet{chang2017splashback} found evidence for the splashback feature
in the weak lensing signal, but their analysis was again done using optically
selected clusters in the Dark energy survey that suffer from the
afore-mentioned issues.

@@

>>>>

\begin{abstract}
We present evidence for the existence of the splashback radius in galaxy
clusters selected using the Sunyaev-Zeldovich effect, a sample unaffected by
systematics related to cluster finding in the optical wavelength range. We show
that the deprojected cross-correlation of galaxy clusters found in the Planck
survey with galaxies detected photometrically in the Pan-STARRS survey, shows a
sharp steepening feature (a logarithmic slope steeper than $-3$), which we
associate with the splashback radius. We infer the three-dimensional splashback
radius for the SZ cluster sample to be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$
$\mpch$, where the cluster sample has an average halo mass $M_{\rm
500c}=3.0\times10^{14}\msunh$ at an average redshift of $z=0.18$.  The inferred
value of the splashback radius appears consistent with the expected location
for dark matter halos in the standard cold dark matter paradigm. However, given
the limited precision of our measurements, we cannot conclusively confirm or
rule out the smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius does not
depend upon the galaxy magnitude for galaxies fainter than $M_i-5\log
h=-19.44$, and is present at a consistent location in galaxy populations
divided by color. The presence of the splashback radius in the star-forming
galaxy population could potentially be used to put lower limits on the
quenching timescales for galaxies. We can marginally rule out the contamination
of the star-forming galaxy sample by quenched galaxies, but the results would
need further verification with deeper datasets.

@@

\textbf{To avoid such optical cluster selection effects, there have been some
previous efforts to use X-ray selected cluster samples to look for the
splashback radius using the weak lensing signal \citep{umetsu2017lensing,
contigiani2018weak}}. However stacking issues and the low signal-to-noise
ratio remains a significant hurdle for both of them. \textbf{The use of weak
lensing data in these studies could get around the biased nature of the galaxy
distribution by directly probing the dark matter distribution. Along those
lines,} \citet{chang2017splashback} found evidence for the splashback feature
in the weak lensing signal, but their analysis was again done using optically
selected clusters in the Dark energy survey that suffer from the
afore-mentioned issues.

@@

In order to avoid optical cluster selection effects, the aim of this paper is
to find observational evidence for the existence of the splashback radius in
galaxy clusters which have not been selected using optical photometry. To move
away from the optical selection we use galaxy clusters selected using the
Sunyaev-Zeldovich effect. We describe the data we use in this section.
